Phyloseq Preprocessing

### 2021-11-05 Rmarkdown/preprocessing/Phyloseq_Preprocessing_211104.Rmd
Added alpha diversity to meta data and downstream work flow

1) Processed QIIME2 inputs in R .rds and .csv formats
2) Phyloseq object with XXX phenotypes

@@

@@

```{r Biocrates BD1 log transformed phyloseq preprocess}
# Load Mary's original memtadata file... that we converted into a csv in the above chunk. 
md <- readRDS("../../data/processed/microbiome/phyloseq_inputs/FL100_mapping_file_complete_highest-seq-reps_nocontrols_min.rds")
rownames(md) <- md[,"Description"]

@@

# Save the phyloseq object
save(PSOlogBioc, file = "../../data/processed/microbiome/phyloseq_objects/PSOlogBioc_211104.RData")
```

>>>>

1) Processed QIIME2 inputs in R .rds and .csv formats
2) Phyloseq object with log transformed Biocrates BD1 data [PSOlogBioc]
3) Phyloseq object with TMAO releated phenotypes [PSOtmao]

@@

Alpha diversity statistics

```{r Load Diversity .qza files}
md <- readRDS("../../data/processed/microbiome/phyloseq_inputs/FL100_mapping_file_complete_highest-seq-reps_nocontrols_min.rds")
rownames(md) <- md[,"Description"]

# Faith PD
FaithPD <- read_qza(file = "../../data/raw/from_MaryKable/core-metrics-results-dada2-tmao/faith_pd_vector.qza", tmp = "Q2RTemp", rm = F)
FaithPD_DF <- FaithPD$data #353, not 355...

# Shannon
shannon <- read_qza(file = "../../data/raw/from_MaryKable/core-metrics-results-dada2-tmao/shannon_vector.qza", tmp = "Q2RTemp", rm = F)
shannon_DF <- shannon$data #353, not 355...

# Observed ASV
Observed <- read_qza(file = "../../data/raw/from_MaryKable/core-metrics-results-dada2-tmao/observed_otus_vector.qza", tmp = "Q2RTemp", rm = F)
Observed_DF <- Observed$data #353, not 355...

# Evenness ASV
evenness <- read_qza(file = "../../data/raw/from_MaryKable/core-metrics-results-dada2-tmao/evenness_vector.qza", tmp = "Q2RTemp", rm = F)
evenness_DF <- evenness$data #353, not 355...

# Merge all of the alpha diversity measures by row names (or a common column name)
l = list(FaithPD_DF, shannon_DF, Observed_DF, evenness_DF)
# Make into one DF
Q2_alpha_DF <- Reduce(merge, lapply(l, function(x) data.frame(x, rn = row.names(x))))

# Now modify (and save downstream)
Q2_alpha_DF$rn <- gsub("[r,x]$","",Q2_alpha_DF$rn)
Q2_alpha_DF$rn <- gsub("\\-.*$","",Q2_alpha_DF$rn)

# Merge with MetaData
metadata2 <- merge(md, Q2_alpha_DF,
                   by.x = 0,
                   by.y = "rn",
                   all.x = TRUE)

rownames(metadata2) <- metadata2[,"Row.names"]
metadata2 <- subset(metadata2, select =-c(Row.names))

# comment out the all.x line in the merge command and run this code; this shows you that 6072 and 7063 don't have alpha stats.
#md2[!(rownames(md2) %in% metadata2$Row.names),] # 6072 and 7063 get left out... no alpha stats for them

# Save
write.csv(metadata2, file = "../../data/processed/microbiome/phyloseq_inputs/FL100_mapping_file_complete_highest-seq-reps_nocontrols_min_withAlphaStatistics_211104.csv")
saveRDS(metadata2, file = "../../data/processed/microbiome/phyloseq_inputs/FL100_mapping_file_complete_highest-seq-reps_nocontrols_min_withAlphaStatistics_211104.rds")

```

@@

```{r Biocrates BD1 log transformed phyloseq preprocess}
# Load Mary's original memtadata file... that we converted into a csv in the above chunk. 
md <- readRDS("../../data/processed/microbiome/phyloseq_inputs/FL100_mapping_file_complete_highest-seq-reps_nocontrols_min_withAlphaStatistics_211104.rds")
rownames(md) <- md[,"Description"]

@@

# Save the phyloseq object
save(PSOlogBioc, file = "../../data/processed/microbiome/phyloseq_objects/PSOlogBioc_211105.RData")
```
